Parsing Tweets into Universal Dependencies

### 2018-03-29 paper.tex
wip

\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
\newcommand{\yicomment}[1]{\textcolor{gray}{[#1 ---\textsc{yi}]}}
\newcommand{\nss}[1]{\textcolor{magenta}{[$_\mathrm{S}^\mathrm{NS}$#1]}}
%\newcommand{\yjcomment}[1]{}
%\newcommand{\nascomment}[1]{}
%\newcommand{\yicomment}[1]{}

@@

\author{Yijia Liu \\
  Harbin Institute of Technology \\
  {\tt oneplus.lau@gmail.com} \\\And
  Yi Zhu \\
  University of Cambridge \\
  {\tt yz568@cam.ac.uk} \\\AND
  Wanxiang Che \quad Bing Qin\\
  Harbin Institute of Technology\\\And
  Nathan Schneider \\
  Georgetown University \\\And
  Noah A. Smith \\
  University of Washington
  }

@@

Universal Dependencies \citep[UD]{NIVRE16.348} were introduced to enable
consistent annotation across different languages. To allow such
consistency, UD was designed to be adaptable to different genres \cite{wang-EtAl:2017:Long6} 
and languages \cite{guo-EtAl:2015:ACL-IJCNLP2,TACL892}. We propose that analyzing
the syntax of tweets can benefit from such adaptability. In this paper,
we introduce a new English tweet treebank of 55,607 tokens that follows the UD
guidelines, but also contends with social media-specific challenges that were not
covered by UD guidelines.\footnote{We developed our treebank independently of a similar effort for Italian tweets \cite{sanguinetti-17}. 
See \nss{section ref} for a comparison.} Our annotation includes 
tokenization, part-of-speech (POS) tags, and (labeled) Universal Dependencies.
We characterize the disagreements among our annotators and find that
consistent annotation is still challenging to deliver even with
the extended guidelines.

@@

Before turning to UD annotations, we (re)annotated the data with 
POS tags, for consistency with other UD efforts,
which adopt the universal POS tagset of \newcite{PETROV12.274}.
% The POS annotation of a word in tweets is generally based on its syntactic distribution, which indicates that the POS of word changes on different context.
% Sticking to the syntactic distribution makes it possible that a `$\heart$' symbol is tagged as a verb as shown in Figure \ref{fig:non-syn-toks}. \yjcomment{Need a better explanation.}
%We follow the UD V2 morphology guideline and use the Universal POS tags~ to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}
In some cases,  non-corresponding tag conflicts arose between the UD English 
treebank conventions \cite[UD\_English;][]{Marneffe2014UniversalSD}\footnote{\url{https://github.com/UniversalDependencies/UD_English}}
and the conventions of \newcite{gimpel-EtAl:2011:ACL-HLT2011}.  %\yicomment{The conflicts
                                %should refer to a {\bf
                                %noncorresponding tag} from that of
                                %gimpel to UDEnglish, cuz they dont
                                %have the same tagset.} 
In these cases, we always
conformed to UD, enabling consistency (e.g., when we exploit the
existing UD\_English treebank in our parser for tweets, \S\ref{sec:parsing}).  For example,  
the nominal URL in Figure \ref{fig:informal-toks} is tagged as {\it
  other} ({\tt X}) and {\tt +} is tagged as {\it symbol} ({\tt SYM})
rather than {\it conjunction} ({\tt CCONJ}).  
% Such violation is resulted by our attempt of conforming the annotation to the {\it universal dependencies - English dependency treebank} (UD\_English), in which all the URLs are tagged as {\it other} and symbolic conjunctors are tagged as {\it symbol}.
% By conforming our annotation to the UD\_English data, hopefully, the tools we built on tweets can benefit from the dataset on broader domain. \yjcomment{need wording}

@@

\paragraph{Comparison to PoSTWITA-UD.} \nss{discuss similarities and differences of conventions to \cite{sanguinetti-17}. 
We expect that the official UD guidelines will eventually adopt standards for these issues so the treebanks can be harmonized.}

@@

\subsubsection{Data Collection}
{\sc Tweebank v2} is built on the original data of {\sc Tweebank v1}
(840 unique tweets, 639/201 for training/test set), along with an 
additional 210 tweets sampled from the POS-tagged dataset of
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and 2,500 tweets sampled
from the Twitter stream from February 2016 to July 2016.\footnote{Data downloaded from \url{https://archive.org/}.}
The latter data source consists of 147.4M English tweets after being
filtered by the {\it lang} attribute in the tweet JSON and {\it
  langid.py}.\footnote{\url{https://github.com/saffsd/langid.py}}

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrr}
		\hline
		& train & dev & test \\
		\hline
		tweets 		&	1,639 		&	709		&	1,201\\
		& +1,000 & +709 & +1,000\\
		tokens 		&	24,753		&	11,742	&	19,112\\
		\hline
	\end{tabular}
\caption{Statistics of {\sc Tweebank v2}.
The second row shows the number of new tweets compared to
{\sc Tweebank v1}.\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets.\nss{What tools did annotators use?}
Newly created annotations are split into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UDPipe v1.2 & 94.7 \\
		\hdashline
		UDPipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} show that the parser\nss{tokenizer?} trained 
on the combination of the UD\_English and Singapore English dialect (Singlish) treebanks
outperforms the one trained only on the dialect treebank.  Preliminary results
showed the same in our case, so 
we follow their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set. % \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} are rule-based systems and were not adapted
to the UD tokenization scheme.
The UDPipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with UDPipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 vs.~20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.\nss{this paragraph needs copyediting}

@@

\paragraph{Learning from exploration.} When we set $\alpha =1$, we
eliminate the oracle from the estimation procedure (for the distilled
model).  This presents an opportunity to learn with \emph{exploration}, by
randomly sampling transitions from the ensemble, found useful
in recent methods for training greedy models that use dynamic oracles
\citep{ballesteros-EtAl:2016:EMNLP2016}.  
We find that this
approach outperforms the  conventional distillation model, coming in
only one point behind the ensemble (last line of Table~\ref{tbl:parse-result}).

@@

>>>>

\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
\newcommand{\yicomment}[1]{\textcolor{gray}{[#1 ---\textsc{yi}]}}
\newcommand{\nss}[1]{\textcolor{magenta}{[$_\mathrm{S}^\mathrm{NS}$#1]}}

@@

\author{Yijia Liu \\
  Harbin Institute of Technology \\
  {\tt yjliu@ir.hit.edu.cn} \\\And
  Yi Zhu \\
  University of Cambridge \\
  {\tt yz568@cam.ac.uk} \\\AND
  Wanxiang Che \quad Bing Qin\\
  Harbin Institute of Technology\\\And
  Nathan Schneider \\
  Georgetown University \\\And
  Noah A. Smith \\
  University of Washington
  }

@@

Universal Dependencies \citep[UD]{NIVRE16.348} were introduced to enable
consistent annotation across different languages. To allow such
consistency, UD was designed to be adaptable to different genres \cite{wang-EtAl:2017:Long6} 
and languages \cite{guo-EtAl:2015:ACL-IJCNLP2,TACL892}. We propose that analyzing
the syntax of tweets can benefit from such adaptability. In this paper,
we introduce a new English tweet treebank of 55,607 tokens that follows the UD
guidelines, but also contends with social media-specific challenges that were not
covered by UD guidelines.\footnote{We developed our treebank independently of a similar effort for Italian tweets \cite{sanguinetti-17}. 
See \S\ref{sec:postwita} for a comparison.} Our annotation includes 
tokenization, part-of-speech (POS) tags, and (labeled) Universal Dependencies.
We characterize the disagreements among our annotators and find that
consistent annotation is still challenging to deliver even with
the extended guidelines.

@@

Before turning to UD annotations, we (re)annotated the data with 
POS tags, for consistency with other UD efforts,
which adopt the universal POS tagset.\footnote{A revised and extended version of \citet{PETROV12.274} with 17 tags.}
% The POS annotation of a word in tweets is generally based on its syntactic distribution, which indicates that the POS of word changes on different context.
% Sticking to the syntactic distribution makes it possible that a `$\heart$' symbol is tagged as a verb as shown in Figure \ref{fig:non-syn-toks}. \yjcomment{Need a better explanation.}
%We follow the UD V2 morphology guideline and use the Universal POS tags~ to tag the tweet tokens. For twitter specific linguistic phenomena, we have set different strategies and try to align with the UD conventions and UD\textunderscore English as close as possible.\footnote{\url{https://github.com/UniversalDependencies/UD\_English}}
In some cases,  non-corresponding tag conflicts arose between the UD English 
treebank conventions \cite[UD\_English;][]{Marneffe2014UniversalSD}\footnote{\url{https://github.com/UniversalDependencies/UD_English}}
and the conventions of \newcite{gimpel-EtAl:2011:ACL-HLT2011}.  %\yicomment{The conflicts
                                %should refer to a {\bf
                                %noncorresponding tag} from that of
                                %gimpel to UDEnglish, cuz they dont
                                %have the same tagset.} 
In these cases, we always
conformed to UD, enabling consistency (e.g., when we exploit the
existing UD\_English treebank in our parser for tweets, \S\ref{sec:parsing}).  For example,  
the nominal URL in Figure \ref{fig:informal-toks} is tagged as {\it
  other} ({\tt X}) and {\tt +} is tagged as {\it symbol} ({\tt SYM})
rather than {\it conjunction} ({\tt CCONJ}).  
% Such violation is resulted by our attempt of conforming the annotation to the {\it universal dependencies - English dependency treebank} (UD\_English), in which all the URLs are tagged as {\it other} and symbolic conjunctors are tagged as {\it symbol}.
% By conforming our annotation to the UD\_English data, hopefully, the tools we built on tweets can benefit from the dataset on broader domain. \yjcomment{need wording}

@@

\subsection{Comparison to PoSTWITA-UD}\label{sec:postwita}
\citet{sanguinetti-17} presented the \textit{PosTWITA-UD} which is
the first Twitter treebank in Universal Dependencies.
In their convention, tokenization tends to preserve the original
tweet content but two special cases, \textit{articulated prepositions}
and \textit{clitic clusters}, are tokenized.
Beyond tokenization, an extra layer of lemmas is annotated
to recover ``what is understandable by a human''.
It's arguable that normalization (the lemmatization in \citet{sanguinetti-17})
on social media data is essential according to \citet{eisenstein:2013:NAACL-HLT}.
In this paper, we don't include normalization and leave it an open question for further study.

In the syntax annotation, tweet specific
constructions are considered by both \citet{sanguinetti-17} and our paper.
We agree on the syntax head in these constructions but differ on
the choice of relations for depicting these constructions.
We use \textit{discourse} for topical hashtag and \citet{sanguinetti-17} used \textit{parataxis}. 
Similar difference happens in referential URLs where we use \textit{list} by following UD\_English and they use \textit{dep}.
The choice of relation on the non-syntactic tokens is questionable
since UD doesn't have specified guideline for this. 
Our choice of \textit{discourse} is inspired by the observation that emoticon is annotated as \textit{discourse}. Non-syntactic tokens can be easily recovered
from our annotation and changing it to other relation in the future development is not difficult.

%Ho \textit{non-syntactic tokens} (like sentiment emoticon, topical hashtag).

\nss{discuss similarities and differences of conventions to \cite{sanguinetti-17}. 
We expect that the official UD guidelines will eventually adopt standards for these issues so the treebanks can be harmonized.}
\yjcomment{i've finished this part.}

@@

\subsubsection{Data Collection}
{\sc Tweebank v2} is built on the original data of {\sc Tweebank v1}
(840 unique tweets, 639/201 for training/test set), along with an 
additional 210 tweets sampled from the POS-tagged dataset of
\newcite{gimpel-EtAl:2011:ACL-HLT2011} and 2,500 tweets sampled
from the Twitter stream from February 2016 to July 2016.\footnote{Data downloaded from \url{https://archive.org/}.}
The latter data source consists of 147.4M English tweets after being
filtered by the {\it lang} attribute in the tweet JSON and {\it
  langid.py}.\footnote{\url{https://github.com/saffsd/langid.py}}
In the same way as \citet{kong-EtAl:2014:EMNLP2014},
reference unit is always the tweet in its entirety
-- which may thus consist of multiple sentences -- not the sentence alone.
Before annotation, we use simple regular expression to anonymize username and URL.

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrrr}
		\hline
		 & \multicolumn{2}{c}{\textsc{Tweebank v1}} & \multicolumn{2}{c}{\textsc{Tweebank v2}} \\
		split & tweets & tokens & tweets & tokens \\
		\hline
		train & 639 & 9,310 & 1,639 & 24,753 \\
		dev & - & - & 709 & 11,742 \\
		test & 201 & 2,839 & 1,201 & 19,112 \\
		\hline
		total & 840 & 12,149 & 3,549 & 55,617 \\
%		\textsc{v1} & tweets & 639 & 0 & 201 & 840 \\
%		& tokens  & 9,310 & 0 & 2,839 & 12149 \\
%		\hdashline
%		\textsc{v2} & tweets & 1,639 & 709 & 1,201 & 3,549 \\
%		 & tokens &	24,753	&	11,742	&	19,112 & 55,617\\
		\hline
	\end{tabular}

%	\begin{tabular}{rrrrrr}
%		\hline
%		\multicolumn{2}{l}{version} & train & dev & test & total \\
%		\hline
%		\textsc{v1} & tweets & 639 & 0 & 201 & 840 \\
%		& tokens  & 9,310 & 0 & 2,839 & 12149 \\
%		\hdashline
%		\textsc{v2} & tweets & 1,639 & 709 & 1,201 & 3,549 \\
%		 & tokens &	24,753	&	11,742	&	19,112 & 55,617\\
%		\hline
%	\end{tabular}
\caption{Statistics of {\sc Tweebank v2} and the comparison with
{\sc Tweebank v1} in quantity.\nss{not obvious how to interpret this: train tokens looks like 1,609+1,000 = 2,609, but maybe it's 609+1,000=1,609? 
I think it would be easier just to show v1 and v2 as separate rows. Also, a column for totals across the splits.}
\yjcomment{i've fix this.}
\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets\nss{What tooks did annotators use?}\yjcomment{fixed}.\footnote{We use arborator \cite{gerdes:2013:W13-37} in our annotation.}
Newly created annotations are split into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UDPipe v1.2 & 94.7 \\
		\hdashline
		UDPipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} show that the parser\nss{tokenizer?}\yjcomment{it's a parsing work.} trained 
on the combination of the UD\_English and Singapore English dialect (Singlish) treebanks
outperforms the one trained only on the dialect treebank.  Preliminary results
showed the same in our case, so 
we follow their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set. % \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} are rule-based systems and were not adapted
to the UD tokenization scheme.
The UDPipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with UDPipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 vs.~20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.\nss{this paragraph needs copyediting}

@@

\paragraph{Learning from exploration.} When we set $\alpha =1$, we
eliminate the oracle from the estimation procedure (for the distilled
model).  This presents an opportunity to learn with \emph{exploration}, by
randomly sampling transitions from the ensemble, found useful
in recent methods for training greedy models that use dynamic oracles
\citet{goldberg-nivre:2012:PAPERS,TACL145,TACL885,ballesteros-EtAl:2016:EMNLP2016}.  
We find that this
approach outperforms the  conventional distillation model, coming in
only one point behind the ensemble (last line of Table~\ref{tbl:parse-result}).

@@

\section*{Acknowledge}
We thank Elizabeth Clark, Lucy Lin, Nelson Liu, Kelvin Luu, Phoebe Mulcaire, 
Hao Peng, Maarten Sap, Chenhao Tan, Sam Thomson,
and the other colleagues in Georgetown University for the annotation efforts in the first round.
We also thank the anonymous  reviewers for their helpful comments and suggestions.
\yjcomment{fonding agency here and anyone else i miss.}
This work was supported by the National Key Basic Research
Program of China via grant 2014CB340503 and the
National Natural Science Foundation of China (NSFC) via
grant 61632011.
